Hydra Battles and Cie

### 2021-07-09 doc/part-hydras.tex
update the doc on ON
minor corrections

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\begin{Coqsrc}
Section Definitions.

@@

Infix "o<" := ON_lt : ON_scope.
Infix "o<=" := ON_le : ON_scope.
\end{Coqsrc}

@@

Compute ON_compare 6 9.
\end{Coqsrc}

@@

\begin{Coqsrc}
Example ex1' : fin 7 o< omega.
Proof. constructor. Qed.

Lemma lt_omega alpha : 
     alpha o< omega <-> exists n:nat,  alpha = fin n.
(* ... *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
\end{Coqsrc}

@@

Definition t := (B * A)%type.
Definition lt : relation t := lexico ltB ltA.
Definition le := clos_refl _ lt.

@@

Lemma compare_reflect alpha beta :
  match (compare alpha beta)
  with
    Lt => lt alpha  beta
  | Eq => alpha = beta
  | Gt => lt beta  alpha
  end.

Global Instance ON_mult : ON lt compare.

@@

The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the last section.

@@

\begin{Coqsrc}
Definition Omega2 := ON_mult Omega Omega.
Existing Instance Omega2.
Definition t := ON_mult.t nat nat.
\end{Coqsrc}

@@

Definition limitb (alpha : t) := 
 match alpha with
 |  (S _, 0) => true
 | _ => false
end.
\end{Coqsrc}

@@

The successor of any ordinal is defined by a simple pattern-matching.
This function is proved to be correct w.r.t. the \texttt{Successor} predicate.

@@

@@

Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma t0_empty (alpha: t 0): False.
Proof.
  destruct alpha.
  destruct x; cbn in i; discriminate.
Qed.

@@

Program Example gamma1 : t 8 := 7.

@@

In order to build an instance of \texttt{OrdinalNotation}, we define a comparison function, by delegation to standard library's  \texttt{Nat.compare}, and prove its correction.

@@

\begin{Coqsrc}
Global Instance sto n : StrictOrder (@lt n).

Global Instance FinOrd (n:nat) : OrdinalNotation (sto n) compare.
Proof.
  split.
  - apply compare_correct.
  - apply lt_wf.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON_correct `(alpha : Schutte_basics.Ord)
     `(OA : @ON A ltA  compareA)
      (iota : A -> Schutte_basics.Ord) :=
  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
                                exists b, iota b = beta;
    On_compare_spec : forall a b:A,
        match compareA a b with
          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
        | Datatypes.Eq => iota a = iota b
        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
        end
}.
\end{Coqsrc}

>>>>

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\label{sect:measure-ON}
\begin{Coqsrc}
Section Definitions.

@@

Infix "o<" := ON_lt : ON_scope.
Infix "o<=" := ON_le : ON_scope.
Infix "o?=" := ON_compare (at level 70) : ON_scope.
\end{Coqsrc}

@@

#[local] Open Scope ON_scope.
Compute 6 o?= 9.
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute  fin 8 o?= omega.
\end{Coqsrc}

\begin{Coqsrc}
Lemma lt_omega alpha : 
     alpha o< omega <-> exists n:nat,  alpha = fin n.
(* ... *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition Zero_limit_succ_dec : ZeroLimitSucc_dec.
Proof.
(* ... *)
Defined.
\end{Coqsrc}

@@

Definition t := (B * A)%type.
Definition lt : relation t := lexico ltB ltA.

@@

#[global] Instance ON_mult: ON lt compare.

@@

The ordinal $\omega^2$ (also called $\phi_0(2)$, see Chap.~\ref{chap:schutte}), is an instance of the multiplication presented in the preceding section.

@@

\begin{Coqsrc}
Instance Omega2: ON _ _ := ON_mult Omega Omega.

Definition t := ON_mult.t nat nat.
\end{Coqsrc}

@@

Definition limitb (alpha : t) : bool := 
 match alpha with
 |  (S _, 0) => true
 | _ => false
 end.

 Definition succb (alpha: t) : bool :=
 match alpha with
  (_, S _) => true
  | _ => false
end.
\end{Coqsrc}

@@

The successor of any ordinal is defined by a simple pattern-matching.

@@

This function is proved to be correct w.r.t. the \texttt{Successor} predicate.

\begin{Coqsrc}
Lemma succ_ok alpha beta : Successor beta alpha <-> beta = succ alpha.
\end{Coqsrc}

@@

@@

\begin{Coqsrc}
Lemma t0_empty (alpha: t 0): False.
Proof.
  destruct alpha; discriminate.
Qed.

@@

\begin{Coqsrc}
Program Example gamma1 : t 8 := 7.

@@

In order to build an instance of \texttt{ON}, we define a comparison function,  and prove its correction.

@@

\begin{Coqsrc}
#[global] Instance sto n : StrictOrder (@lt n).

#[global] Instance comp n: Comparable (@lt n) compare.

#[global] Instance FinOrd n : ON (@lt n) compare. 
Proof.
  split.
  - exact (comp n).
  - apply lt_wf.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON_correct `(alpha : Ord)
     `(OA : @ON A ltA compareA)
      (iota : A -> Ord) :=
  { ON_correct_inj : forall a, lt (iota a) alpha;
    ON_correct_onto : forall beta, lt beta alpha ->
                                exists b, iota b = beta;
    On_compare_spec : forall a b:A,
        match compareA a b with
          Datatypes.Lt => lt (iota a) (iota b)
        | Datatypes.Eq => iota a = iota b
        | Datatypes.Gt => lt (iota b) (iota a)
        end
  }.
\end{Coqsrc}
